### 2020-10-17 inverse_problems/C2.tex

\section{Classical regularisation theory}
\begin{definition}
    Let $A \in \LLL(\XXX, \YYY)$ and 
\end{definition}