### 2020-10-19 inverse_problems/C2.tex

%\section{Classical regularisation theory}
%\begin{definition}
%    Let $A \in
%\end{definition}