### 2020-11-10 inverse_problems/C4.tex
I'm still alive

\section{Convex Duality}
We are looking for a convergence rate
\[
D_\JJ^\Rm{symm}(u_\delta, u_\JJ\D) \leq \psi(\delta) \qquad \psi \colon \RR_+ \to \RR_+, \quad \lim_{\delta\to0}\psi(\delta) = 0. 
\]

We will assume for simplicity that $\JJ$ is absolutely one-homogeneous.  We define
\[
E(y) \ceq \frac12 \norm{y - f}_\YY^2, \quad F(u) \ceq \alpha \JJ(u), 
\]
so that $u_\delta \in \argmin_{u \in \XX} E(u) + F(Au)$.

\begin{lemma}
	Let $X$ be a Banach space and $\phi(x) = \frac12\norm{x}_X^2$. Then the convex conjugate of $\phi$ is
	\[
	\phi^*(\xi) = \frac12 \norm{\xi}_{X^*}^2. 
	\]
\end{lemma}

\begin{proof}
	We have
	\[
	\phi^*(\xi) = \sup_{x \in X} \ang{\xi, x} - \frac12 \norm{x}^2 \leq \sup_{x \in X} \norm{x} \norm{\xi} - \frac12 \norm{x}^2.
	\]
	The right-hand is a quadratic function in $\norm{x}$ with maximum $\frac12\norm{\xi}^2$. The inequality is in fact equality (\TODO show) which proves the result. 
\end{proof}

Now we compute
\[
E^*(\eta) = \ang{\eta, f} + \frac12 \norm{\eta}_{\YY^*}, \quad F^*(p) = \chi_{\pt \JJ(0)} (p/\alpha), 
\]
which gives the dual problem
\[
\sup_{\eta \in \YY^*} - \ang{\eta, f} - \frac12\norm{\eta}^2 - \chi_{\pt \JJ(0)} \qty(-\frac{A^*\eta}{\alpha}),
\]
which is equivalent to
\[
\sup_{\mu \in \YY^*, A^*\mu \in \pt \JJ(0)} \alpha \qty(\ang{\mu, f} - \frac\alpha2 \norm{\mu}^2)
\]

We need to