### 2020-11-16 modern_statistical_methods/msm_ES3.tex

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[margin=3cm]{geometry}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools, amsmath, amssymb, amsthm, mdframed, bbm, graphicx, float, physics, xcolor, cleveref}

\hypersetup{
    colorlinks   = true, %Colours links instead of ugly boxes
    urlcolor     = blue, %Colour for external hyperlinks
    linkcolor    = blue, %Colour of internal links
    citecolor   = red %Colour of citations
}

% Definition of numbered environments.
% Usage: \begin{theorem} ... \end{theorem}
% Remark has no numbering.
\theoremstyle{plain}
\newtheorem{question}{Question}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

% Question with box around it
\newenvironment{qbox}{\begin{mdframed}\begin{question}}{\end{question}\end{mdframed}}

% A proof with "solution" instead of "proof" and no QED symbol
\newenvironment{solution}{\begin{proof}[Solution]\renewcommand\qedsymbol{}}{\end{proof}}

% Some renewed commands
\renewcommand{\vec}{\mathbf}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\theta}{\vartheta}
\renewcommand{\phi}{\varphi}

% Frequently used math alphabets
\newcommand{\Bb}{\mathbb}
\newcommand{\Cal}{\mathcal}
\newcommand{\Bf}{\mathbf}
\newcommand{\Rm}{\mathrm}

% Frequently used letters in the blackboard alphabet
\newcommand{\CC}{\Bb C}
\newcommand{\NN}{\Bb N}
\newcommand{\PP}{\Bb P}
\newcommand{\QQ}{\Bb Q}
\newcommand{\RR}{\Bb R}
\newcommand{\EE}{\Bb E}

\newcommand\XX{\Cal X}
\newcommand\HH{\Cal H}

% Usage: \ang{...} is equivalent to \langle ... \rangle, while \ang*{...} is equivalent to \left\langle ... \right\rangle
% For other delimiters: use \qty from the physics package (i.e., \qty(...))
\DeclarePairedDelimiter{\ang}{\langle}{\rangle}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

% Frequently used commands
\newcommand{\T}{^\top} % Matrix transpose A\T
\newcommand{\C}{^\complement} % Set complement A\C
\newcommand\ceq\coloneqq % Definitions :=
\newcommand\pow{\Cal P} % Power sets
\newcommand\eps\epsilon
\newcommand\ind{\mathbbm 1} % Blackboard 1 for indicator functions
\newcommand\restr{\mathord\restriction}
\newcommand\TODO{{\color{red} TODO: }}
\renewcommand\P{^\perp}

% Functions that appear frequently
\DeclareMathOperator{\sign}{sgn}
\DeclareMathOperator{\Int}{Int}
\DeclareMathOperator{\Span}{Span}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand\Nul{\Cal N}
\newcommand\Ran{\Cal R}

\title{Modern Statistical Methods --- Example Sheet 3} % subject
\author{Lucas Riedstra}
%\date{} % date

\begin{document}
\maketitle
\begin{mdframed}
	In all of the below, assume that any design matrices $X$ are $n \times p$ and have their columns centred and then scaled to have $\ell^2$-norm $\sqrt n$, and that any responses $Y \in \RR^n$ are centred. 
\end{mdframed}
\begin{question}
	When proving the theorems on the prediction error of the Lasso, we started with the so-called basic inequality that
	\[
	\frac1{2n} \norm{X(\beta^0 - \hat\beta)}_2^2 \leq \frac1n \eps\T X(\hat\beta - \beta^0) + \lambda \norm{\beta^0}_1 - \lambda \norm{\hat\beta}_1. 
	\]
	Show that in fact we can improve this to 
		\[
	\frac1{n} \norm{X(\beta^0 - \hat\beta)}_2^2 \leq \frac1n \eps\T X(\hat\beta - \beta^0) + \lambda \norm{\beta^0}_1 - \lambda \norm{\hat\beta}_1. 
	\]
\end{question}

\begin{proof}
	By the KKT conditions, we there exists $\hat\nu \in \RR^p$ with $\norm{\hat\nu}_\infty \leq 1$ and $\ang{\hat\nu, \hat\beta} = \norm{\beta}_1$ such that 
	\begin{align*}
		\frac1n X\T (Y - X\hat\beta) &= \lambda\hat\nu \\
		\frac1n X\T (X(\beta^0 - \hat\beta) + \eps) &= \lambda\hat\nu \\
		\frac1n X\T X(\beta^0 - \hat\beta) &= - \frac1n X\T\eps + \lambda\hat\nu \\
		\frac1n (\beta^0 - \hat\beta)\T X\T X(\beta^0 - \hat\beta) &= -\frac1n (\beta^0 - \hat\beta)\T X\T\eps + \lambda (\beta^0 - \hat\beta)\T \hat\nu \\
		\frac1n \norm{X(\beta^0 -  \hat\beta)}_2^2 &= \frac1n \eps\T X(\hat\beta - \beta^0) + \lambda \ang{\hat\nu, \beta^0} - \lambda \norm{\hat\beta}_1,
		\end{align*}
	and now plugging in $\ang{\beta^0, \hat\nu} \leq \norm{\beta^0}_1 \norm{\hat\nu}_\infty \leq \norm{\beta^0}_1$ yields the result. 
\end{proof}

\begin{question}
	Under the assumptions of Theorem 23 on the prediction and estimation properties of the Lasso under a compatibility condition, show that, with probability $1 - 2p^{-(A^2/8 - 1)}$, we have
	\[
	\frac1n \norm{X(\beta^0 - \hat\beta)}_2^2 \leq \frac{9A^2 \log(p)}{4\phi^2} \frac{\sigma^2 s}{n}. 
	\]
\end{question}
\end{document}